

Microcontroller operating systems and frameworks typically assume that a single,
monolithic application will run on an embedded system.
Traditionally,
optimizing for power, squeezing applications into minimal code and memory
allocations, and lacking hardware support
% In fact, historically, power draw, memory and code sizes, and lack of hardware support
have
constrained microcontroller applications to be single-function.
Newer hardware, however, has changed this paradigm.
The microcontroller is growing up and can now support
a secure, trusted kernel and
multiple, isolated, concurrent, and dynamically-loaded applications,
all while operating on the power budgets that originally made this device
class feasible.
While the hardware support is present, the software ecosystem to capitalize on these
advances is lagging behind. To remedy this, we propose Tock, a new embedded operating system design
that builds on established operating system
principles while adapting to the embedded system environment. Tock exploits
memory protection units, advancements in modern systems programming languages,
and the event driven nature of embedded applications to allow a core kernel,
device specific drivers, and untrusted applications to coexist on a single
microcontroller. This new operating system will allow embedded devices to
mature beyond program-once, deploy-once systems and into re-usable, ubiquitous,
and reliable computing platforms.




% This is no
% longer the case with modern hardware. Modern microcontrollers, like the Atmel
% SAM4L (an ARM Cortex-M4), provides over six times more SRAM and over five times
% larger flash than the TI~MSP430 that powered the TelosB motes while maintaining
% similar power draws (90~{\uA}/MHz active and 3~\uA at sleep).
% Simultaneously embedded products are becoming a development platforms and an
% application ecosystems for products like the Pebble watch. Additionally, a set
% of modular embedded devices is emerging. Devices like SimBand, Wzzard and
% ThinkingThings have a core part to which additional modules can be attached to
% extend the functionality. These modules can be developed by the third parties
% and have own microcontroller with runtime environment.
% The operating systems community should leverage advancements in hardware,
% programming languages as well as our experience from the Web, and other
% application rich ecosystems, to build the next generation of embedded operating
% systems.

