\section{Challenges}

\subsection{Using a Safe Programming Language}

\begin{enumerate}
  \item Accessing globals is fundamentally part of operating systems development
    (e.g. hardware memory registers), but access to global variables is not
    thread-safe and therefore considered ``unsafe'' in languages like Rust.
    While there is prior art wrapping an unsafe core with an safe interface,
    it's not clear where to draw the line -- which code should be trusted with
    access to unsafe features?

  \item Certain common constructs in high-level programming languages like
    closures often rely on dynamic allocation outside the stack. Embedded
    operating systems usually avoid allocating memory dynamically since memory
    is usually limited in embedded environments and the kernel cannot rely on
    user intervention to recover from fatal runtime errors---instead embedded
    operating systems prefer to rely on compile-time guarantees that memory
    requirements are met.

  \item Other concerns
\end{enumerate}

\subsection{Doing Without Virtual Memory}

\begin{enumerate}
  \item Swapping due to memory use
  \item Allocating application memory. Specifically, virtual memory affords the
    operating system the ability to allocate pages dynamically to an application
    as it requires them (for example, dynamic stack growth in high-level
    languages). This is harder without virtual memory since memory that appears
    continuous to applications {\emph must} be continuous in physical memory.
    Moreover, relocating memory would require changing pointers in application
    code and data structures.
\end{enumerate}

\subsection{Heterogenous Platforms}

Platforms with multiple microprocessors are not like multi-core desktop
operating systems. They typically do not share caches or even main memory.
Often microcontrollers, while from a similar family (like the cortex-m) have
different manufaturer specific features and configurations or are different
architecture versions. As a result, we cannot trivially move work from one
processor to another since that would require relocating memory pointers and
code may simply not be portable.
